---
title: "L1_md"
author: "Álvaro Pérez"
date: "2025-01-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Getting Started

## Console commands

R console is to use it as a calculator

```{r}
1 + 1
```

You can break up mathematical expressions over multiple lines

```{r}
## let's break it over multiple lines
1 + 2 + 3 + 4 + 5 + 6 +
    7 + 8 + 9 +
    10
```

Text inside quotes is called a string. You can break up text over multiple lines; R waits for a close quote before processing it.

```{r}
"Good afternoon"
```

## Objects

An object in R:

-   contains only letters, numbers, full stops, and underscores

-   starts with a letter or a full stop and a letter

-   distinguishes uppercase and lowercase letters

Use the assignment operator\<-\` to assign the value on the right to the object named on the left.

```{r}
## use the assignment operator '<-'
## R stores the number in the object
x <- 5
```

## Environment

Anytime you assign something to a new object, R creates a new entry in the global environment. Objects in the global environment exist until you end your session; then they disappear forever

```{r}
ls()            # print the objects in the global environment
rm("x")         # remove the object named x from the global environment
rm(list = ls()) # clear out the global environment
```

## Whitespace

R mostly ignores whitespace: spaces, tabs, and line breaks. This means that you can use whitespace to help you organise your code.

```{r}
# a and b are identical
a <- list(ctl = "Control Condition", exp1 = "Experimental Condition 1", exp2 = "Experimental Condition 2")

# but b is much easier to read
b <- list(ctl  = "Control Condition", 
          exp1 = "Experimental Condition 1", 
          exp2 = "Experimental Condition 2")
```

When you see \> at the beginning of a line, that means R is waiting for you to start a new command.

## Function syntax

If you want 10 random numbers from a normal distribution with mean of 0 and standard deviation, you can just use the defaults.

```{r}
rnorm(10)
```

If you want 10 numbers from a normal distribution with a mean of 100:

```{r}
rnorm(10, 100)
```

This would be an equivalent but less efficient way of calling the function:

```{r}
rnorm(n = 10, mean = 100)
```

## Add-on packages

Remember: **Install ≠ Load**

```{r}
install.packages("ggExtra")
```

### Loading a package

```{r}
library(ggExtra)
```

## Organising a project

```{r}
# load add-on packages
library(tidyverse)

# set object ----
n <- 100

# simulate data ----
data <- data.frame(
  id = 1:n,
  dv = c(rnorm(n/2, 0), rnorm(n/2, 1)),
  condition = rep(c("A", "B"), each = n/2)
)

# plot data ----
ggplot(data, aes(condition, dv)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.25, 
               aes(fill = condition),
               show.legend = FALSE)

# save plot ----
ggsave("sim_data.png", width = 8, height = 6)
```

## Working directory

```{r}
setwd("/Users/alvaroperezlopez/Desktop/Curso R")
```
